# Friends by Chance

Before we start, let's execute [a helper script](https://github.com/j-rohrer/marginal-psych/blob/main/scripts/load.R) that loads the necessary dependencies.

```{r}
#| warning: false
#| message: false
source(here::here("scripts/load.R"))
```

## Overview

This document contains example 3, in which we analyze the effect of being seated next to each other on friendships, using Bayesian multilevel models.
Data can be downloaded from the [OSF](https://osf.io/wt7nh){target="_blank"} but are also included in the downloadable replication package. 

Every day experience—and previous research—suggests that being spatially close to others can result in friendships. 
But does spatial proximity also lead to friendships for people who are quite different from each other? 
We re-analyze data from a large field experiment conducted in 3rd to 8th grade classrooms in rural Hungary previously reported in [Rohrer et al. (2021)](https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0255097){target="_blank"}. 
Proximity was experimentally manipulated by randomizing each classrooms’ seating chart at the beginning of the school semester; thus, students randomly ended up next to each other (deskmate = 1) or not (deskmate = 0). 
At the end of the semester, students listed up to five best friends from their classroom, which allows us to determine which pairs of students had formed friendships (friendship = 1). 
Additionally, we know students’ gender and grade point average before the experiment (GPA), which allows us to investigate whether proximity also “works” for girls seated next to boys (who are quite unlikely to befriend each other at that age) or students with discrepant levels of academic achievement.



## Read and clean the data

```{r load_fit}
# Read the data
dat <- read.csv(here("data/deskmates.csv"))

# Arbitrary subset of 60 classes to make model-fitting less tedious
dat <- dat[dat$class_id %in% unique(dat$class_id)[1:60], ]

# Keep complete cases
dat <- dat[complete.cases(dat),]


# Label focal variables
dat <- transform(dat,
    gender_combination = factor(girl_match, label = c("Boys", "Mixed", "Girls")),
    deskmate = factor(deskmate, label = c("Different desk", "Same desk")))

# Rename to match labels in manuscript
dat <- dat |>
  rename(GPA_average = mean_gpa,
         GPA_difference = diff_gpa,
         friendship = friend,
         classroom = class_id,
         student1 = s1,
         student2 = s2)


# Display the first few rows of data
head(dat)
```
## Bayesian multilevel regression model

Our model here ends up a bit more complex due to the nested structure of the data, and we use the brms package [(Bürkner, 2018)](https://journal.r-project.org/archive/2018/RJ-2018-017/RJ-2018-017.pdf){target="_blank"} which allows us to fit multilevel models in a highly flexible manner. 
Since we now fit a Bayesian model (relying on the default priors provided by the package), marginaleffects will return credible intervals rather than confidence intervals. 
Our unit of observation is pairs of students, which are nested within students and classrooms. 
For each pair, we know:

* whether they are deskmates
* their gender_combination (both boys, one girl and one boy, both girls)
* their GPA_average (i.e., the average across both students) and their absolute GPA_difference (i.e., the discrepancy between both students)
* whether they report a friendship at the end of the experiment or not

Here, we fit and save the model to avoid refitting repeatedly.

```{r, eval = FALSE}
mod <- brm(
    friendship ~ deskmate + gender_combination + deskmate:gender_combination +
        GPA_average + GPA_difference + GPA_average:deskmate + GPA_difference:deskmate +
        (1 | classroom) + (1 | mm(student1, student2)),
    family = bernoulli(link = "logit"),
    data = dat,
    seed = 12345,
    cores = 4
)

# Save model fit
saveRDS(mod, file = here("fits/deskmates.rds"))
```

```{r}
# Load the fitted model
mod <- readRDS(here("fits/deskmates.rds"))

# Look at the model coefficients
parameters(mod)
```

## Interpretation with `marginaleffects`

### What is the average treatment effect of being a deskmate on the log-odds of a friendship (link scale)?

In principle, we could evaluate this model on the log-odds scale on which the coefficients are estimated.

```{r}
avg_comparisons(mod, variables = "deskmate", type = "link")
```

However, log-odds of friendship are not a particularly intuitive unit, and so instead we may want to switch to the scale of the outcome (friendships)—which is the default behavior of marginaleffects:

### What is the average treatment effect of being a deskmate on the probability of a friendship?

```{r}
# Predicted probabilities
avg_predictions(mod, by = "deskmate")

# Average Treatment Effect
ate <- avg_comparisons(mod, variables = "deskmate")
ate
```

This returns the average effect of the intervention in percentage points.

### Comparing our effect to the fast friends procedure

One may want to compare the effect of sitting next to each other to other interventions meant to foster friendships. 
For example, one staple of psychological research is the fast friends procedure in which two participants are paired up and then take turns answering questions that escalate in the degree of self-disclosure involved, from mild (“Would you like to be famous? In what way?”) to severe (“When did you last cry in front of another person?”). 
[Echols and Ivanich (2021)](https://onlinelibrary.wiley.com/doi/abs/10.1111/jora.12622){target="_blank"} implemented such a procedure in US middle school students and found that those who underwent the intervention in three sessions over three months were 10 percentage points more likely to become friends. 
This seems very close to the 11 percentage point effect we observed in our analysis.

Would it be justified to conclude that the effects are practically the same?

In a Frequentist framework, this would be a use case for an equivalence test, like the Two One Sample Test, or TOST (Lakens et al., 2018). 
All the functions in the marginaleffects package have an equivalence argument that can execute a TOST for any of the quantities that the package estimates: predictions, comparisons, slopes, and even for complex quantities derived for non-linear hypothesis tests. 
Since the model that we estimated above is Bayesian, marginaleffects will not conduct a TOST by default. 
Instead, the equivalence argument will quantify the size of the posterior density in a region of practical equivalence (ROPE; Kruschke, 2018; Makowski et al., 2019).

To do this, we first need to define a range around the estimated effect size of the fast friends procedure. 
This range must be based on substantive and application-specific considerations. 
For instance, the analyst could decide that if the effect of deskmates falls within ± a quarter of the fast friends effect, they will be considered “equivalent” to each other. In that case, the analyst would specify a ROPE of [0.075; 0.125]. 
Now, we can use the equivalence argument to calculate how likely it is that the effect of sitting next to each other falls into this ROPE. 

```{r}
avg_comparisons(mod, variables = "deskmate", 
                       equivalence = c(0.075, 0.125)) 

```

### What is the average treatment effect of being a deskmate on the probability of a friendship, at desks with different gender compositions?

Having looked at the average effect of the intervention, we still do not yet know whether being seated next to each other also “works” for dissimilar students. 
Here, we will keep evaluating effects on the outcome scale, which we consider most intuitive.
First, we can separately calculate average effects depending on gender_combination: 


```{r}
avg_comparisons(mod, variables = "deskmate", by = "gender_combination")

```

#### Pairwise comparisons, e.g.: Is the effect of sharing a desk stronger for students on girls-only desks, relative to the effect of sharing a desk for students on mixed-gender desks?

We can easily compare all three average effects against each other in a pairwise manner:

```{r}
avg_comparisons(mod,
                variables = "deskmate",
                by = "gender_combination",
                hypothesis = ~pairwise)

```


#### Is the effect of sharing a desk stronger for students on gender-matched desks, relative to the effect of sharing a desk for students on mixed-gender desks?

We may also want to compare gender-matched dyads (pairs of girls, pairs of boys) with gender-mismatched dyads (pairs of a girl and a boy), (b_Two girls + b_Two boys)/2 = b_One girl one boy. 
This can be achieved by using a different hypothesis argument (being mindful of the order in which the groups are listed in our output):

```{r}
avg_comparisons(mod,
                variables = "deskmate",
                by = "gender_combination",
                hypothesis = "(b1 + b3)/2 = b2")

```

#### Plotting the effects by gender

We can also visualize these different effects:

```{r}
# Plot the predicted probabilities
plot_predictions(mod, by = c("gender_combination", "deskmate"))

# Plot the effects of the intervention in percentage points
plot_comparisons(mod, variables = "deskmate", by = "gender_combination")

```


### Causal Moderation and All-Else-Equal Claims

In the previous section, we used the avg_comparisons() function and its by argument to compute the average effect of deskmate within each subgroup of gender_combinations. 
Under the hood, marginaleffects started by computing unit-level estimates in the full dataset, and then aggregated those estimates within each level of the moderator variable. 
This approach makes sense if our goal is simply to establish whether the effect (or association) is stronger in some subsets of the population. 
Indeed, our analysis showed that the positive effect of deskmate on the probability of striking a friendship was much stronger in matched gender pairs than in mixed gender pairs. 
Unfortunately, the estimates reported above cannot tell us why the effects differ between subgroups. 
They cannot establish that the effect changes because of the moderator, because the strata defined by the moderator may also vary in other variables that could explain the difference in effects.  

Researchers often do have the more ambitious research goal of establishing that some moderator causally affects the effect of interest (for the distinction between the non-causal and the causal “type” of moderation, see Bansak, 2021; VanderWeele, 2009; for an overview of different terminologies used for this distinction, see Rohrer and Arslan, 2021). 
In that context, researchers must pay special attention to the distribution of potential confounders in the various moderator strata. 
If such confounders of the moderation exist, it is important to conduct a counterfactual analysis that holds their distribution constant across moderator strata.

Crucially, only confounders should be held constant.
Holding constant variables that are potential mediators or colliders may actually increase bias ([Rohrer, 2018](https://journals.sagepub.com/doi/full/10.1177/2515245917745629){target="_blank"}; [Wysocki et al., 2022](https://journals.sagepub.com/doi/10.1177/25152459221095823){target="_blank"}).

First, let’s say we are interested in whether gender_combination causally interacts with deskmate. 
Recall that our model also contains GPA_difference. 
Should we control for GPA_differences between dyads to isolate the causal effect of gender_combination on the effects of deskmate?
Different gender_combinations may indeed vary in their GPA_difference, but notice that these differences are most plausibly causally downstream of gender_combination. 
For gender-mismatched dyads, we may observe larger differences in GPA because gender affects GPA (Gender combination → GPA difference → Effect of being seated next to each other). 
Thus, GPA_difference is a potential mediator of gender_combination and should not be conditioned on when evaluating whether gender_combination causally affects the effects of being seated next to each other. 
This vindicates our earlier analysis.

Now, let’s imagine we were interested in whether GPA_differences causally affect the effects deskmate. 
For this other estimand, gender_combination is a confounder (GPA difference ← Gender combination → Effect of being seated next to each other). 
Thus, to evaluate the causal interaction between GPA_difference and deskmate, we do want to control for gender_combination. 
If differences in grades affect the effect of the intervention, then we should be able to observe an association between differences in grades and the effects of the intervention among dyads of the same type (e.g., among all female dyads).

So, does the effect of being seated next to each other (deskmate) vary by how strong their academic achievement diverges (GPA_difference), controlling for the effects of gender match (gender_combination)? 
We can answer this question by conducting comparisons on new hypothetical datasets in which we set GPA_difference to particular values (e.g., ± 1 SD) but keep the other variables (including gender_match) as is. 

#### What is the average treatment effect of sharing a desk for students with high and low differences in GPA, accounting for gender?


```{r}
avg_comparisons(mod,
                variables = "deskmate", # cause of interest
                by = "GPA_difference", # variable to split by
                newdata = datagrid(GPA_difference = c(mean(dat$GPA_difference, na.rm = TRUE) - sd(dat$GPA_difference, na.rm = TRUE), 
                                                mean(dat$GPA_difference, na.rm = TRUE) + sd(dat$GPA_difference, na.rm = TRUE)),
                                   grid_type = "counterfactual"))
# the newdata argument is crucial here: we apply the model to new data which represents a counterfactual constrast
# between a world in which everybody's GPA difference is set to 1 SD below the mean
# and a world in which everybody's GPA difference is set to 1 SD above the mean.

# Compare the two estimates
avg_comparisons(mod,
                variables = "deskmate", # cause of interest
                by = "GPA_difference", # variable to split by
                newdata = datagrid(GPA_difference = c(mean(dat$GPA_difference, na.rm = TRUE) - sd(dat$GPA_difference, na.rm = TRUE), 
                                                mean(dat$GPA_difference, na.rm = TRUE) + sd(dat$GPA_difference, na.rm = TRUE)),
                                   grid_type = "counterfactual"),
                hypothesis = "b1 = b2")

```


