# German Socio-Economic Panel Study

Before we start, let's execute [a helper script](https://github.com/j-rohrer/marginal-psych/blob/main/scripts/load.R) that loads the necessary dependencies.

```{r}
source(here::here("scripts/load.R"))
```

## Overview

This document contains the analysis code underlying the running example in the manuscript section on target quantities, as well as the code that generates Figure 1.
Annotations were later added manually using PowerPoint.

## Read the data and fit the model

```{r load_fit}

source(here::here("scripts/load.R"))

## We analyze freely available practice data from the German Socio-Economic Panel Study (SOEP).
## The data can be downloaded from the SOEP website (https://www.diw.de/en/diw_01.c.836543.en/soep_practice_dataset.html);
## We are using the English version (DOI: 10.5684/soep.practice.v36).

## Read data
soep <- read_dta(here("data/practice_en/practice_dataset_eng.dta"))
head(soep)

## Fit simple model that predicts life satisfaction
mod <- lm(
  lebensz_org ~ sex * bs(alter, df = 3) * bs(einkommenj1, df = 3),
  data = soep[soep$alter < 60, ]
)
## this model is fairly flexible insofar that it flexibly models the effects of age and income
## and also allows for all interactions between the variables
## we limit ourselves to people under the age of 60


## let's take a look at the coefficients to confirm they are not easily interpretable
summary(mod)
```

## Calculating various quantities

```{r interpretation}

## Prediction: life satisfaction of a 35 year old woman who earns 20,000 euro
predictions(mod, newdata = data.frame(sex = 1, alter = 35, einkommenj1 = 20000))

## Store value for plotting purposes
prediction_20000 <- predictions(
  mod,
  newdata = data.frame(sex = 1, alter = 35, einkommenj1 = 20000)
)$estimate


## Prediction: what if she earned twice as much?
predictions(mod, newdata = data.frame(sex = 1, alter = 35, einkommenj1 = 40000))

## Store value for plotting purposes
prediction_40000 <- predictions(
  mod,
  newdata = data.frame(sex = 1, alter = 35, einkommenj1 = 40000)
)$estimate


## Comparison: Compare these two predictions
comparisons(
  mod,
  newdata = data.frame(sex = 1, alter = 35, einkommenj1 = 20000),
  variables = list("einkommenj1" = c(20000, 40000))
)


## Slope: How much does life satisfaction increase with income at 20000 euro
slopes(
  mod,
  newdata = data.frame(sex = 1, alter = 35, einkommenj1 = 20000),
  variables = "einkommenj1"
)

## Store value for plotting
slope_20000 <- slopes(
  mod,
  newdata = data.frame(sex = 1, alter = 35, einkommenj1 = 20000),
  variables = "einkommenj1"
)$estimate

## Let's translate this into life satisfaction points per 1,000 Euro
slopes(
  mod,
  newdata = data.frame(sex = 1, alter = 35, einkommenj1 = 20000),
  variables = "einkommenj1"
)$estimate *
  1000
```

## Generate Figure 1

```{r figure}

## We *could* directly do this with the help of marginaleffects using the following:
plot_predictions(
  mod,
  newdata = data.frame(
    sex = 1,
    alter = 35,
    einkommenj1 = seq(from = 0, to = 80000, by = 10)
  ),
  by = "einkommenj1"
)

## However, let's take things slowly and do this manually
## Just for ease of plot-optimization

## First step: Generate a data frame containing the same 35-year-old woman under different incomes
hypothetical_woman <- data.frame(
  sex = 1,
  alter = 35,
  einkommenj1 = seq(from = 0, to = 80000, by = 10)
)

## Let's generate her life satisfactions (predicted life satisfaction)
hypothetical_predictions <- predictions(mod, newdata = hypothetical_woman)
head(hypothetical_predictions)

## Add the life satisfaction to the data we are going to plot
hypothetical_woman$lebensz_org <- hypothetical_predictions$estimate


## Let's plot!

## Just for customization: x-axis range over which the slope should be plotted
slope_range <- c(0, 60000)

## Color scheme
color_prediction <- "#0072B2"
color_comparison <- "#56B4E9"
color_slope <- "#009E73"
color_neutral <- "#BBBBBB"


ggplot(data = hypothetical_woman, aes(x = einkommenj1, y = lebensz_org)) +
  # Prediction line
  geom_line() +
  # Comparison 20,000 vs 40,000 (connected with  segments that form a slope triangle)
  geom_segment(
    x = 20000,
    xend = 40000,
    y = prediction_20000,
    color = color_neutral,
    linetype = "dashed"
  ) +
  geom_segment(
    x = 40000,
    y = prediction_20000,
    yend = prediction_40000,
    color = color_comparison,
    linetype = "solid"
  ) +
  # Slope line at an income of 20,000
  geom_segment(
    x = slope_range[1],
    xend = slope_range[2],
    y = prediction_20000 - (20000 - slope_range[1]) * slope_20000,
    yend = prediction_20000 + (slope_range[2] - 20000) * slope_20000,
    color = color_slope
  ) +
  # Predictions at 20,000 and 40,000
  geom_point(x = 20000, y = prediction_20000, color = color_prediction) +
  geom_point(x = 40000, y = prediction_40000, color = color_prediction) +
  # Layout, make it look nicer
  xlab("Annual gross income") +
  ylab("Predicted life satisfaction") +
  coord_cartesian(ylim = c(7, 8.25)) +
  scale_x_continuous(
    labels = label_dollar(prefix = "", suffix = "€", big.mark = ",")
  )

# ggsave(here("plots/predictions_comparisons_slopes.png"), width = 4, height = 3)
```





